---
title: "BeHo tube test - initial statistical analyses"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data preparation and transformation

I prepare the tube test data for the analyses, which involved data transformation etc. I've added all relevant commands for you to run my commands. This document is however still incomplete. 

```{r cars}
library(tidyverse)
library(reshape2)
library(tidyr)
library(effects) 
library(ggplot2) 
library(sjPlot)

setwd("~/Documents/PhD 2020/BeHo:ZiBA study/Data analysis/Tube test data analysis")
tube_test_wins = read.csv("R_input_files/Mice-full-overview.csv", header = TRUE)

#Placing NAs for dead mice (C8 mouse 3065, C16 mouse 3183, C7 mouse 2730, C3 mouse 3083, C1 mouse 4176)
tube_test_wins[76, c(32:45,48:49)] = NA
tube_test_wins[38, c(32:45,48:49)] = NA
tube_test_wins[32, c(32:45,48:49)] = NA
tube_test_wins[14, c(32:45,48:49)] = NA
tube_test_wins[4,c(28:45,47:49)] = NA

#Removing irrelevant columns
tube_test_80 <- tube_test_wins [,-c(5,8,10,15:17,21:23, 27, 31, 35:37, 39, 42:50)]

#Renaming columns without the many dots
names(tube_test_80)[2] <- "Arrive_weight"
names(tube_test_80)[5] <- "Sex"
names(tube_test_80)[6] <- "Exp_group"

#Transform the dataframe from a wide to a long format
tube_test_80 <- melt(tube_test_80, id.vars = c("Name","Arrive_weight", "Phenotype", "Crate", "Sex", "Exp_group","Cage.ID"))

#Renaming the columns created by melt()
names(tube_test_80)[8] <- "Treatments"
names(tube_test_80)[9] <- "Wins_per_rep"
names(tube_test_80)[7] <- "Cage_ID"
names(tube_test_80)[4] <- "Crate_ID"

#Add a new column with the replicate info and then rename the rows of column Treatments to only the treatments 
Replicates <- c(rep("R1",80), rep("R2",80), rep("R3",80),rep("R4",80),rep(c(rep("R1",80), rep("R2",80), rep("R3",80)),4),rep("R3",80), rep("R4",80), rep("R5",80))
#add the column "Replicates" to the dataframe tube_test_80
tube_test_80 <- cbind(tube_test_80,Replicates)
#Changing the name of the factors created after transforming the dataframe. There are 19 in total corresponding to the no of reps in total.
levels(tube_test_80$Treatments) <- c("OPT","OPT","OPT","OPT","HEAT","HEAT","HEAT","COLD","COLD","COLD","DIET","DIET","DIET","ANTI","ANTI","ANTI","FMT","FMT","FMT")
levels(tube_test_80$Exp_group) <- c("Treatment", "Control") #does the glm or lm model then know how to differ from the experimental and control animals tube test wins? Ask A & O this.

#Move the no of wins column (all of it) to the first column
colnames(tube_test_80)
tube_test_80 <- tube_test_80[, c(9,8,10,1:7)]

#Creating new factors to start statistics. A factor instead of num, chr or int account for variation between the "catagories"
tube_test_80 <- transform(tube_test_80,Replicates=as.factor(Replicates),
                            Phenotype=as.factor(Phenotype),
                            Crate=as.factor(Crate_ID),
                            Sex=as.factor(Sex),
                            Exp_group=as.factor(Exp_group),
                            Cage_ID=as.factor(Cage_ID))
str(tube_test_80)
head(tube_test_80)

#Remove mice that are not among the 40 I'm working with:
#I have mice from cage 3-6 (males) and 10-13 (females)
cages_to_remove <- c("01M","02M","07M","08M","09F","14F","15F","16F")
condition <- tube_test_80$Cage_ID %in% cages_to_remove
tube_test_40 <- tube_test_80[!condition, ]
# Remove rows with missing values (NA) in any column if needing the df like this later
cleaned_tube_test_40 <- drop_na(tube_test_40)

str(tube_test_40)
head(tube_test_40)

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
summary(cars)
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
